Estimating the genome-wide contribution of selection to temporal allele frequency change

### 2019-10-19 manuscript/manuscript.tex
updates to ms

% - change supp fig ordering
% - sig impact
% -ref table 1
% fix equation G(t) in Fig 1

@@

\begin{table}
  \centering
  \footnotesize
  \centering
  \begin{tabular}{l c c c c c c}
    \textbf{Study} & \textbf{Species} & \textbf{Selection} & \textbf{Replicates} & \textbf{Pop. Size} & \textbf{Gens.} & \textbf{Timepoints} \\ 
  \hline
    \textcite{Kelly2019-dc}  & \emph{D. simulans} & lab adaptation & 3 & $\sim$1100 & 14 & 2 \\ \hline
    \textcite{Barghi2019-qy} & \emph{D. simulans} & lab adaptation & 10 & $\sim$1000 & 60 & 7 \\ \hline
    \textcite{Castro2019-uk} & \emph{M. musculus} & \begin{tabular}{c}tibiae length\\control\end{tabular} & \begin{tabular}{c}2\\1\end{tabular} &\begin{tabular}{c}32\\28\end{tabular}  & 20 & 2 \\
 \hline
  \end{tabular}
  \caption{A summary of the main selection studies we analyzed.}
\end{table}

@@

% \Section{Materials and Methods}

% \subsection{Datasets Analyzed}

% \subsection{Partitioning Unique and Shared Selection Effects}

% We can partition the allele frequency change between the two timepoints (20
% generations) for a Longshanks line as \vb{$\Delta p_{t,\mathrm{LS1}} =
% \Delta_{_{D}} p_{t,\mathrm{LS1}} + \Delta_{_{U}} p_{t,\mathrm{LS1}} +
% \Delta_{_S} p_{t,\mathrm{LS}}$} where these  terms in are the decomposition are
% the allele frequency change due to drift in \vb{Longshanks replicate 1
%   ($\Delta_{_D} p_{t,\mathrm{LS1}}$), selection unique to the LS1 replicate
%   ($\Delta_{_U} p_{t,\mathrm{LS1}}$), and selection response shared between the
% two Longshanks replicates ($\Delta_{_S} p_{t,\mathrm{LS}}$) respectively (and
% similarly for the Longshanks two line, LS2)}. By construction we will assume
% that each of these terms are uncorrelated within replicates, and that only the
% shared term covaries between the replicates. Assuming that we can approximate
% the contribution of genetic drift in the LS lines the variance in allele
% frequency change observed in the control, i.e. \vb{$\var(\Delta
% p_{t,\mathrm{Ctrl}}) = \var(\Delta_{_{D}}
% p_{t,\mathrm{LS2}})=\var(\Delta_{_{D}} p_{t,\mathrm{LS2}})$}, then we can
% estimate the increase in variance in allele frequency change due to selection
% as $(\var(\Delta p_{t,\mathrm{LS1}}) + \var(\Delta p_{t,\mathrm{LS2}}))/2 -
% \var(\Delta p_{t,\mathrm{Ctrl}})$ and the shared effect of selection across
% selected lines as $\cov(\Delta p_{t,\mathrm{LS1}}, \Delta p_{t,\mathrm{LS2}})$.

% \subsection{Variance and Covariance Estimates}

% We calculate the variance-covariance matrix of allele frequency changes using a
% Python software package we have written, available at
% \url{http://github.com/vsbuffalo/cvtk}. Our method uses the sampling depth and
% number of diploid individuals sampled to correct the bias in the variance
% estimates and a bias that occurs in covariance estimates between adjacent
% timepoints (see Supplementary Material Sections \ref{supp:ind-depth-var-corr},
% \ref{supp:cov-corr}, and \ref{supp:matrix-correction} for mathematical details
% of these estimators).

% \subsection{Block Bootstrap}

>>>>

% TODO
% - sig impact
% - ref table 1
% - more correction plots?

@@

\begin{table}
  \centering
  \footnotesize
  \centering
  \begin{tabular}{l c c c c c c}
    \textbf{Study} & \textbf{Species} & \textbf{Selection} & \textbf{Replicates} & \textbf{Pop. Size} & \textbf{Gens.} & \textbf{Timepoints} \\ 
  \hline
    \textcite{Kelly2019-dc}  & \emph{D. simulans} & lab adaptation & 3 & $\sim$1100 & 14 & 2 \\ \hline
    \textcite{Barghi2019-qy} & \emph{D. simulans} & lab adaptation & 10 & $\sim$1000 & 60 & 7 \\ \hline
    \textcite{Castro2019-uk} & \emph{M. musculus} & \begin{tabular}{c}tibiae length\\control\end{tabular} & \begin{tabular}{c}2\\1\end{tabular} &\begin{tabular}{c}32\\28\end{tabular}  & 17 & 2 \\
 \hline
  \end{tabular}
  \caption{A summary of the main selection studies we analyzed.}
\end{table}

@@

\Section{Materials and Methods}

\subsection{Datasets Analyzed}

We used available genomic data data from four studies: pooled population
resequencing (pool-Seq) data from \textcite{Barghi2019-qy},
\textcite{Kelly2019-dc}, \textcite{Bergland2014-ij}, and
\textcite{Castro2019-uk}. In all cases, we used the variants kept after the
filtering criteria of the original studies.

\subsection{Variance and Covariance Estimates}

To remove systematic covariances in allele frequency change caused by tracking
the reference or minor allele, we randomly choose which allele's frequency to
track for each locus. Then, we calculate the variance-covariance matrix of
allele frequency changes using a Python software package we have written,
available at \url{http://github.com/vsbuffalo/cvtk}. This simultaneously
calculates temporal variances and covariances, and replicate covariances and
uses the sampling depth and number of diploid individuals to correct for bias
in the variance estimates and a bias that occurs in covariance estimates
between adjacent timepoints due to shared sampling noise (see Supplementary
Material Sections \ref{supp:ind-depth-var-corr}, \ref{supp:cov-corr}, and
\ref{supp:matrix-correction} for mathematical details of these estimators). We
assess that our bias correction procedure is working adequately through a
series of diagnostic plots that ensure that the procedure removes the
relationship between sampling depth and uncorrected variance and covariances
(Supplementary Figure \ref{suppfig:bergland-correction}, XXX).

\subsection{Estimating $G(t)$ and the Convergence Correlation}

\subsection{Block Bootstrap}

\subsection{Partitioning Unique and Shared Selection Effects}

We can partition the allele frequency change between the two timepoints (20
generations) for a Longshanks line as \vb{$\Delta p_{t,\mathrm{LS1}} =
\Delta_{_{D}} p_{t,\mathrm{LS1}} + \Delta_{_{U}} p_{t,\mathrm{LS1}} +
\Delta_{_S} p_{t,\mathrm{LS}}$} where these  terms in are the decomposition are
the allele frequency change due to drift in \vb{Longshanks replicate 1
  ($\Delta_{_D} p_{t,\mathrm{LS1}}$), selection unique to the LS1 replicate
  ($\Delta_{_U} p_{t,\mathrm{LS1}}$), and selection response shared between the
two Longshanks replicates ($\Delta_{_S} p_{t,\mathrm{LS}}$) respectively (and
similarly for the Longshanks two line, LS2)}. By construction we will assume
that each of these terms are uncorrelated within replicates, and that only the
shared term covaries between the replicates. Assuming that we can approximate
the contribution of genetic drift in the LS lines the variance in allele
frequency change observed in the control, i.e. \vb{$\var(\Delta
p_{t,\mathrm{Ctrl}}) = \var(\Delta_{_{D}}
p_{t,\mathrm{LS2}})=\var(\Delta_{_{D}} p_{t,\mathrm{LS2}})$}, then we can
estimate the increase in variance in allele frequency change due to selection
as $(\var(\Delta p_{t,\mathrm{LS1}}) + \var(\Delta p_{t,\mathrm{LS2}}))/2 -
\var(\Delta p_{t,\mathrm{Ctrl}})$ and the shared effect of selection across
selected lines as $\cov(\Delta p_{t,\mathrm{LS1}}, \Delta p_{t,\mathrm{LS2}})$.
